---
title: "Data Wrangling - Part II"
subtitle: "StatPREP R Workshops"
author: "Hailey Watkins"
output: html_document
---

```{r setup, include=FALSE}
# Don't forget to load your libraries
library(tidyverse)
library(ggplot2movies)
library(ggformula)
```

```{r}
# This loads the data from the ggplot2movies package
data(movies)
```

# Deriving information

So far, you have learned "verbs" for isolating data, `select`, `filter`, `arrange`. Now, we will learn verbs for deriving information, `summarize`, and `mutate`. There is an additional verb, `group_by`, which is useful in conjunction with these verbs, but not as much on its own. 

## summarize

Summarize is a many-to-one or many-to-few operation. It summarizes the dataset down to fewer numbers. 

Let's say we want to know the average budget of the movies in the `movies` dataset. The way to compute that is with summarize,

```{r}
movies %>%
  summarize(mean(budget))
```

Unfortunately, there are many missing (or Not Available) values in `budget`, so we need to remove those in our pipeline.

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(mean(budget))
```

Notice that this pipeline does not change the `movies` dataset in our environment. The NA values are still present in the data, and the entire dataset is just how it started.  We just `summarize`d the data down to a single number in our Console or RMarkdown document. 

If we want, we can add additional statistics to our `summarize` command,

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(mean(budget), max(budget))
```

We can also name the statistics we compute, so we could refer to them later.

```{r}
movies %>%
  drop_na(budget) %>%
  summarize(avg_budget = mean(budget), max_budget = max(budget))
```
# Question 1
### Your turn

Now, you try it! 

- What year was the earliest movie in the dataset released?  
- What is the average length of movies? 
- What is the standard deviation of length?

You may want to look at the "data transformation with dplyr" cheatsheet to review some possible summary functions. 

```{r eval=TRUE}
movies_lin71<-movies %>%
  summarize(mean(length),sd(length),min(year))
movies_lin71
```


Films produced in 1893 are the oldest on this list. Movies are typically 82 minutes and 34 seconds long. Despite this, there is a standard deviation of roughly 35 seconds and 44 minutes. 

# Question 2

## group_by

The `group_by` function groups cases by common values. 

This is useful if want to compute summary statistics by group. For example, we may be interested in whether action movies are longer than non-action movies. 

We already know how to compute the average length of movies overall. What we need to add to our pipeline is a `group_by`

```{r eval=TRUE}
movies_ln84<-movies %>%
  group_by(Action) %>%
  summarize(mean(length),sd(length),min(year))
  # move your summarize code from the last exercise here!
movies_ln84
```

Which movies are longer, action movies or non-action movies? 

Generally speaking, action films are longer than non-action films, with average durations of around 100 minutes as opposed to 81 minutes. The difference between the two is also evident in the standard deviation, which is lower for action films and is closer to 100 minutes, whereas non-action films fluctuate more in run duration.

# Question 3
### Your turn

Investigate the movies classified as Short. Are they actually shorter than other movies? 

- What is the average length? 
- What is the minimum length?

```{r eval=TRUE}
movies_ln99<-movies %>%
  group_by(Short) %>%
  summarize(mean(length),min(length))
movies_ln99
```

The average length of a non-short film is approximately 95 minutes, but the average length of a short film is nearly 14 minutes. It is interesting to note that the minimum duration of a film is one minute, which is the same for both short and non-short films.

## mutate

The `mutate` function creates new variables (columns) in the dataset. Typically, we use it to make new variables from old ones. 

For example, we may be interested in the length of movies in hours, rather than in minutes. 

```{r}
movies <- movies %>%
  mutate(length_hrs = length / 60)
```


Let's read this code carefully. 

- what does the `<-` do?
- why are we writing `/60`?
- what does the `=` do?
- why did we use the name `length_hrs`?

# Question 4
### Your turn

The variables `r1`, `r2`, ..., `r10` give you the percent of people who rated that movie a 1, 2, ..., 10. We would like to know the percent of people who rated the movie a 4 or less (i.e., who rated it a 1, 2, 3, or 4). Create a new variable called `per_low` that computes that number.

```{r eval=TRUE}
movies_ln128 <- movies %>%
  mutate(per_low = r1+r2+r3+r4)
movies_ln128
```

# Question 5
## Putting it together

Now that we have a bunch of wrangling skills, let's put them together. 

- pull out the movies that had a high percent of low ratings (Hint: what command will you use to do this?)
- pull out movies that had a high percent of low ratings *and* more than 5 votes. 
- how many of these bad movies were there?

To accomplish that last task, we need a specialty function, `n`. Here's how it works

```{r}
movies %>%
  summarize(n())
```

- what is the average length of movies in the dataset, in hours? 
- can you find a second way to compute the average length, by flipping the order of commands (i.e., `summarize` first and then `mutate`)?

```{r}
average_length_hours <- movies %>%
summarize(average_length = mean(length)) %>%
mutate(average_length_hours_summarize = average_length / 60)
average_length_hours_flipped <- movies %>%
mutate(length_hours = length / 60) %>%
summarize(average_length_hours_mutate = mean(length_hours))
average_length_hours
average_length_hours_flipped
```
